---
title: TidyTuesday - Horror movie ratings
author: Felix MIL
date: '2019-10-27'
slug: tidytuesday-horror-movie-ratings
categories: []
tags: []
description: ''
externalLink: ''
series: []
---



```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE, 
                      warning=F, 
                      fig.align = "center", 
                      out.width = "60%", 
                      message=F)
```

## Intro

This post is a submission for Tidy Tuesday 2019 W43: [Horror Movies Ratings](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-10-22)

### Questions

  1. When horror movies became an important part of the movie industry ? My intuition is that horror movies became important (money-wise) in the 2000's and something huge in the 2010's, is it true ? Is it still a growing part of the industry ?
  2. I'm feeling like Northern America (especially the USA) is the most productive source of horror movies. Is it true ? Are the best movies produced in the USA ?
  3. Is there any 'classical' settings in horror movies plots ? As I want to train my Natural Language Processing skills, this seems like an interesting and challenging question.
  4. Is there any frequent location to shoot horror movies ? 

## Preparation

### Loading libraries

```{r}
library(tidyverse)
library(lubridate)
library(forcats)
library(countrycode)
library(ggdark)
```

### Getting the data
```{r}
horror_movies <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-22/horror_movies.csv")
```

### Cleaning & summarising data

The provided data dictionnay is the following:

| variable            | class       | description                           |
| :------------------ | :---------- | :------------------------------------ |
| title               | character   | Title of the movie                    |
| genres              | character   | Movie Generes                         |
| release_date        | character   | Movie release date - day-month-year   |
| release_country     | character   | Release country                       |
| movie_rating        | character   | MPAA Rating                           |
| review_rating       | double      | Movie rating (0 - 10)                 |
| movie_run_time      | character   | Movie run time (minutes)              |
| plot                | character   | Short plot description (raw text)     |
| cast                | character   | Cast                                  |
| language            | character   | Language                              |
| filming_locations   | character   | Filming location                      |
| budget              | character   | Budget (US Dollars)                   |


Let's take a look at the actual data.

```{r}
glimpse(horror_movies)
```


#### `release_date`

Temporal data is crucial to see the trends of Horror movies over time. In the first rows of the dataset, the format seems to be `%d-%b-%y` so `lubridate::dmy()` should be appropriate to parse release dates.

```{r}
  horror_movies %>%
  transmute(release_date,
            release_date_asDate = dmy(release_date))
```

seems to be ok, except some release date only contains year, let's take care of them with `lubridate::parse_date_time()` which allow to pass several potentiel date formats.


```{r}
horror_movies_clean <- 
  horror_movies %>%
  mutate(release_date = parse_date_time(release_date, orders = c("dmy", "%y")))
```

Now we can get the time covered by the dataset:

  - First date : `r min(horror_movies_clean$release_date)`
  - Last date : `r max(horror_movies_clean$release_date)`
  - Duration covered : `r as.duration(interval(min(horror_movies_clean$release_date), max(horror_movies_clean$release_date)))`

Obviously, this dataset is not appropriate to observe trend on a large time scale. Question 1 cannot completely be answered. But we'll still be able to see the latest trends.

#### `release_country`

How many different countries are present in the dataset ?

```{r}
n_distinct(horror_movies$release_country)
```
What are they ?
```{r}
horror_movies %>%
  distinct(release_country) %>%
  arrange(release_country)
```

Apparently, there are no duplicated countries (potentially caused by typos or different denomination, for exemple UK & United Kingdom)

Add a variable `release_contient`
```{r}
horror_movies_clean <- 
  horror_movies_clean %>%
  mutate(release_continent=countrycode(release_country,
                                       origin = "country.name",
                                       destination = 'continent'))
```

#### `review_rating`

The dataset dictionnay indicates this variable is a double between 0 and 10.

```{r}
summary(horror_movies$review_rating)
```

Only `r round(summary(horror_movies$review_rating)["NA's"]/nrow(horror_movies),2)`% of movies are not reviewed

#### `budget`
  #TODO Install [fixer](https://rdrr.io/cran/fixerapi/)
  #TODO create a https://fixer.io/signup/free account
  #TODO Parse currency symbol to 'official' currency strings (€ -> 'EUR')
  #TODO Pass date to get robust dollar consersions with [`fixer_convert`](https://rdrr.io/cran/fixerapi/man/fixer_convert.html)
  
  #TODO OR keep only the ones with $ notation

```{r}
horror_movies_clean <- 
  horror_movies_clean %>%
  mutate(budget_clean=as.numeric(str_remove(str_remove_all(budget,","), "\\$")))
```

## Observations

### Number of movies per year
```{r}
horror_movies_clean %>%
  ggplot(aes(x = as.factor(year(release_date)))) +
  geom_bar(aes(fill=fct_infreq(as.factor(release_continent))),color='white')+
  scale_fill_brewer(type = 'qual',palette = 3)
```

### Repartition of horror movies throughout the month of the years

```{r}
horror_movies_clean %>%
  ggplot(aes(x= month(release_date,label = T,locale="en_US")))+
  geom_bar(aes(fill=release_continent))+
  facet_grid(fct_rev(as.factor(year(release_date)))~.)+
  scale_fill_brewer(type = 'qual',palette = 3)+
  labs(y='# of horror movies', 
       x='month',
       fill='continent:')
```


### Number of movies per country (top 20)
```{r}
horror_movies_clean %>%
  ggplot(aes(fct_lump(fct_infreq(release_country),20)))+
  geom_bar()+
  theme(axis.text.x = element_text(angle=60, hjust = 1))
```

### Distribution of review ratings
```{r}
horror_movies_clean %>%
  ggplot(aes(review_rating))+
  geom_bar(binwidth = 0.2, color='white')
```

```{r}
horror_movies_clean %>%
  ggplot(aes(y=review_rating,x=release_continent))+
  geom_boxplot(outlier.color = NA, fill=NA,width=0.4)+
  geom_jitter(width = 0.1, alpha=0.2, color='firebrick')+
  
  dark_theme_gray()+
  labs(y='Reviews',
       x='Continent')
```

```{r}
horror_movies_clean %>%
  ggplot(aes(y=review_rating,x=release_continent))+
  geom_jitter(width = 0.1, alpha=0.2, color='firebrick')+
  geom_boxplot(outlier.color = NA, fill=NA,width=0.4)+
  dark_theme_gray()+
  labs(y='Reviews',
       x='Continent')
```



### Budgets

```{r}
horror_movies_clean %>%
filter(!is.na(review_rating),
        budget_clean > 0,
       release_continent=='Americas') %>%
  ggplot(aes(budget_clean,review_rating))+
  geom_point(aes(color=release_continent))+
  scale_color_brewer(type='qual',palette=3)+
  coord_flip()
```

```{r}
horror_movies_clean %>%
  group_by(release_continent, m=month(release_date, label = T, locale='EN_US')) %>%
  transmute(med_budget = median(budget_clean,na.rm = T)) %>%
  ggplot(aes(m, med_budget, color=release_continent))+
  geom_point()+
  dark_theme_grey()+
  labs(y='Median Budget', x='Month',color='Continent')+
  scale_y_continuous()
```

## Take Away

More movies are releases in October but budgets for theses movies are generally lower. Reviews are still higher ?


```{r}
horror_movies_clean %>%
  filter(!is.na(budget_clean)) %>%
  ggplot(aes(month(release_date,label = T,locale = 'EN_US'))) +
  geom_bar(aes(fill=month(release_date,label = T,locale = 'EN_US')))+
  dark_theme_gray()+
  scale_fill_viridis_d(option='Plasma',direction = -1,begin=0.1, end = 0.5)+
  guides(fill=FALSE)+
  labs(x='Months', y='# of movies')
```

```{r}
horror_movies_clean %>%
  filter(!is.na(budget_clean)) %>%
  mutate(m = month(release_date,label = T,locale = 'EN_US')) %>%
  ggplot(aes(m,review_rating)) +
  geom_violin(aes(color=m),trim = TRUE)+
  geom_dotplot(
    aes(color= m,
        fill = m),
    binaxis = "y", 
    stackdir = "center",
    dotsize = 0.4
  )+
  dark_theme_gray()+
  scale_fill_viridis_d(option='Plasma',direction = -1,begin=0.1, end = 0.5)+
  scale_color_viridis_d(option='Plasma',direction = -1,begin=0.1,end = 0.5)+
  guides(fill=FALSE, color = FALSE)+
  geom_pointrange(stat = "summary",
    fun.y = median, 
               fun.ymin = function(z) { quantile(z,0.25) }, 
                fun.ymax =function(z) {quantile(z,0.75)},
               color='firebrick')+
  labs(x='Months',y='Reviews')
```


```{r}
horror_movies_clean %>%
  filter(!is.na(budget_clean)) %>%
  mutate(m = month(release_date,label = T,locale = 'EN_US')) %>%
  ggplot() +
  ggridges::geom_density_ridges(aes(x = budget_clean,y=fct_rev(as.factor(m)), fill=m), color='black')+
  scale_x_log10()+
  dark_theme_gray()+
  scale_fill_viridis_d(option='Plasma',direction = -1,begin=0.1, end = 0.5)+
  scale_color_viridis_d(option='Plasma',direction = -1,begin=0.1,end = 0.5)+
  guides(fill=FALSE, color = FALSE)+
  labs(x='log(Budgets)',y='Months')
```


