---
title: TidyTuesday - CRAN code
author: Felix MIL
date: '2019-11-12'
slug: tidytuesday-cran-code
categories: []
tags: []
description: ''
externalLink: ''
series: []
---



```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE, 
                      warning=F, 
                      fig.align = "center", 
                      message=F)
```

## Intro

This post is a submission for Tidy Tuesday 2019 W46: [CRAN code](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-11-12)


## Getting the data

### Data dictionnary

|variable |class     |description |
|:--------|:---------|:-----------|
|file     |double    | Number of files |
|language |character | Programming Language |
|blank    |double    | Blank Lines |
|comment  |double    | Commented Lines |
|code     |double    | Lines of Code |
|pkg_name |character | Package Name |
|version  |character | Package Version |

### A glimpse on data
```{r}
cran_code <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-11-12/loc_cran_packages.csv")
```

```{r}
head(cran_code)
```

```{r}
tibble::glimpse(cran_code)
```

## Question

**Which are the most and less documented package available on CRAN ?**

## Preparation

### Loading libraries

```{r}
library(tidyverse)
library(lubridate)
library(forcats)
library(cranlogs)
```

### Measures

```{r}
cran_code_out <- 
  cran_code %>%
  mutate(pkg_name = as.factor(pkg_name)) %>%
  group_by(pkg_name) %>%
  mutate(comm_code_ratio =  sum(code, na.rm = T)/sum(comment,na.rm = T)) %>%
  filter(comm_code_ratio != Inf) %>%
  select(pkg_name, comm_code_ratio) %>%
  ungroup() %>%
  unique()

cran_code_out <- 
  cran_code_out %>%
  mutate(c = cut(comm_code_ratio, breaks = c(seq(0,20,0.5))))
  
  

mean(cran_code_out$comm_code_ratio)


# sd(cran_code_out$comm_code_ratio)


# cran_code_out <- 
#   cran_code_out %>%
#   mutate(q= cut(comm_code_ratio, breaks=c(quantile(cran_code_out$comm_code_ratio, probs = seq(0, 1, by = 0.20)))))
```



## Visualisation

### 

```{r,fig.asp=0.5}
cran_code_out %>% 
  ggplot(aes(c)) +
  geom_bar()+
  theme(axis.text.x = element_text(angle=30))+
  scale_x_discrete(labels = c(seq(0,19.5,0.5), '20+'))+
  labs(x='Comment per code line', y='count')
```


```{r,fig.asp=0.5}
cran_code_out %>%
  top_n(n = 25,wt = comm_code_ratio) %>%
  ggplot(aes(fct_reorder(pkg_name,comm_code_ratio),comm_code_ratio))+
  geom_col() +
  coord_flip()+
  labs(y='Comment line per code line', x='package',title = 'Top 25 packages with most commented R code')
```


```{r}
cran_code_out %>%
  top_n(n = 25,wt = 1/comm_code_ratio) %>%
  ggplot(aes(fct_reorder(pkg_name,-comm_code_ratio),comm_code_ratio))+
  geom_col() +
  coord_flip()+
    labs(y='Comment line per code line', x='package', 'Bot 25 packages with less commented R code')
```


```{r}
top_down <- 
  cran_top_downloads(when = c( "last-month"), count = 100)

cran_code_out %>% 
  filter(pkg_name %in% top_down$package) %>%
  ggplot(aes(c)) +
  geom_bar()+
  theme(axis.text.x = element_text(angle=30))+
    scale_x_discrete(labels = c(seq(0,19.5,0.5), '20+'))+
  labs(x='Comment lines per code line', y='count', title='Number of comment lines per R code line for the 100 most downloaded packages')
```

