---
title: 'Code Advent 2019'
author: 'Felix MIL'
date: '2019-12-01'
slug: code-advent-2019
categories: []
tags: []
description: ''
externalLink: ''
series: []
---



```{r setup, include=F}
library(reticulate)
knitr::opts_chunk$set(echo = TRUE, 
                      warning=F, 
                      fig.align = "center", 
                      out.width = "60%", 
                      message=F)
use_python("/Applications/miniconda/miniconda3/bin/python")
```


```{r}
library(here)
```

# Motivations

# --- Day 1: The Tyranny of the Rocket Equation ---

## part 1

```
Santa has become stranded at the edge of the Solar System while delivering presents to other planets! To accurately calculate his position in space, safely align his warp drive, and return to Earth in time to save Christmas, he eds > you to bring him measurements from fifty stars.

Collect stars by solving puzzles. Two puzzles will be made available on each day in the Advent calendar; the second puzzle is unlocked when you complete the first. Each puzzle grants one star. Good luck!

The Elves quickly load you into a spacecraft and prepare to launch.

At the first Go / No Go poll, every Elf is Go until the Fuel Counter-Upper. They haven't determined the amount of fuel required yet.

Fuel required to launch a given module is based on its mass. Specifically, to find the fuel required for a module, take its mass, divide by three, round down, and subtract 2.

For example:

For a mass of 12, divide by 3 and round down to get 4, then subtract 2 to get 2.
For a mass of 14, dividing by 3 and rounding down still yields 4, so the fuel required is also 2.
For a mass of 1969, the fuel required is 654.
For a mass of 100756, the fuel required is 33583.
The Fuel Counter-Upper needs to know the total fuel requirement. To find it, individually calculate the fuel needed for the mass of each module (your puzzle input), then add together all the fuel values.
```

**Problem: What is the sum of the fuel requirements for all of the modules on your spacecraft?**

The fuel calculation formula is defined as: ' Fuel required to launch a given module is based on its mass. Specifically, to find the fuel required for a module, > take its mass, divide by three, round down, and subtract 2.' 

In R it would be:

```{r}
compute_fuel <- function(mass) {
  fuel <- floor(mass / 3) - 2
  if (fuel < 0) {
    fuel = 0
  }
  return(fuel)
}
```

Now, let's test this function:
```{r}
test_mass <- c(12,14,1969, 100756)

compute_fuel(test_mass)
```
The test is passed ! 

Now, let's compute the total fuel requirement for or spacecraft.

All modules masses are given as input : `r paste(import(here('resources','data','input.txt'))$V1[1:5],'...')`

```{r}
modules_mass <- as.numeric(readLines(here('resources','data','input.txt')))

total_fuel <- 0
for (m in modules_mass) {
  module_fuel <- compute_fuel(m)
  total_fuel <-  total_fuel + module_fuel
}
```

Total fuel needed for our spacecraft is `r total_fuel`

## part 2

```
--- Part Two ---
During the second Go / No Go poll, the Elf in charge of the Rocket Equation Double-Checker stops the launch sequence. Apparently, you forgot to include additional fuel for the fuel you just added.

Fuel itself requires fuel just like a module - take its mass, divide by three, round down, and subtract 2. However, that fuel also requires fuel, and that fuel requires fuel, and so on. Any mass that would require negative fuel should instead be treated as if it requires zero fuel; the remaining mass, if any, is instead handled by wishing really hard, which has no mass and is outside the scope of this calculation.

So, for each module mass, calculate its fuel and add it to the total. Then, treat the fuel amount you just calculated as the input mass and repeat the process, continuing until a fuel requirement is zero or negative. For example:

A module of mass 14 requires 2 fuel. This fuel requires no further fuel (2 divided by 3 and rounded down is 0, which would call for a negative fuel), so the total fuel required is still just 2.
At first, a module of mass 1969 requires 654 fuel. Then, this fuel requires 216 more fuel (654 / 3 - 2). 216 then requires 70 more fuel, which requires 21 fuel, which requires 5 fuel, which requires no further fuel. So, the total fuel required for a module of mass 1969 is 654 + 216 + 70 + 21 + 5 = 966.
The fuel required by a module of mass 100756 and its fuel is: 33583 + 11192 + 3728 + 1240 + 411 + 135 + 43 + 12 + 2 = 50346.
```

**Problem: What is the sum of the fuel requirements for all of the modules on your spacecraft when also taking into account the mass of the added fuel? (Calculate the fuel requirements for each module separately, then add them all up at the end.)**

The R function:
```{r}
compute_fuel_fuel <-  function(mass) {
  fuel <- compute_fuel(mass)
  fuel_fuel <- compute_fuel(fuel)
  remaining <- fuel_fuel
  while (remaining != 0) {
    remaining <- compute_fuel(remaining)
    fuel_fuel <- fuel_fuel + remaining
  }
  return(fuel + fuel_fuel)
}
```

Test:
```{r}
compute_fuel_fuel(test_mass)
```
```{r}
total_fuel <- 0
for (m in modules_mass) {
  module_fuel <- compute_fuel_fuel(m)
  total_fuel <-  total_fuel + module_fuel
}
```

Now, total fuel requirement is `r total_fuel`. That's a lot Santa ! Please, use solar panels !

# --- Day 2: 1202 Program Alarm ---

## Part I

instruction are too long, but are available [here](https://adventofcode.com/2019/day/2)

**Problem: Once you have a working computer, the first step is to restore the gravity assist program (your puzzle input) to the "1202 program alarm" state it had just before the last computer caught fire. To do this, before running the program, replace position 1 with the value 12 and replace position 2 with the value 2. What value is left at position 0 after the program halts? **


The R function
```{r}
program_alarm <- function(gap) {
  
  op_index = seq(1, length(gap),4)
  
  for (op_i in op_index) {
    op <-  gap[op_i]
    if (op == 1) {
      gap[gap[op_i + 3]+1]<- gap[gap[op_i + 1]+1] + gap[gap[op_i + 2]+1]
    } else if (op == 2) {
      gap[gap[op_i + 3]+1] <- gap[gap[op_i + 1]+1] * gap[gap[op_i + 2]+1]
    } else if (op == 99)
      return(gap)
  }
}
```


Test
  - 1,0,0,0,99 becomes 2,0,0,0,99 (1 + 1 = 2).
  - 2,3,0,3,99 becomes 2,3,0,6,99 (3 * 2 = 6).
  - 2,4,4,5,99,0 becomes 2,4,4,5,99,9801 (99 * 99 = 9801).
  - 1,1,1,4,99,5,6,0,99 becomes 30,1,1,4,2,5,6,0,99.
  
```{r}
program_alarm(c(1,0,0,0,99))
program_alarm(c(2,3,0,3,99))
program_alarm(c(2,4,4,5,99,0))
program_alarm(c(1,1,1,4,99,5,6,0,99))

```


Input:
```{r}
input <- c(1,12,2,3,1,1,2,3,1,3,4,3,1,5,0,3,2,13,1,19,1,10,19,23,1,6,23,27,1,5,27,31,1,10,31,35,2,10,35,39,1,39,5,43,2,43,6,47,2,9,47,51,1,51,5,55,1,5,55,59,2,10,59,63,1,5,63,67,1,67,10,71,2,6,71,75,2,6,75,79,1,5,79,83,2,6,83,87,2,13,87,91,1,91,6,95,2,13,95,99,1,99,5,103,2,103,10,107,1,9,107,111,1,111,6,115,1,115,2,119,1,119,10,0,99,2,14,0,0)
```

```{r}
program_alarm(input)
```

## part 2

**Problem: Find the input noun and verb that cause the program to produce the output 19690720. What is 100 * noun + verb? (For example, if noun=12 and verb=2, the answer would be 1202.)**

```{r}

for (noun in 0:99) {
  for (verb in 0:99) {
    input[2] = noun
    input[3] = verb
    out <- program_alarm(input)[1]
    if (out == 19690720) {
      print(paste('noun:', noun))
      print(paste('verb:',verb))
      print(paste('answer is:', 100 * noun + verb))
    }
  }
  
}
```


# --- Day 3: Crossed Wires ---

instructions are [here](https://adventofcode.com/2019/day/3)

**problem: What is the Manhattan distance from the central port to the closest intersection?**

Today, we'll try to solve that problem using python (thanks to reticulate)

```{python}
import re

cable_1 = "R1001,D890,R317,U322,L481,D597,L997,U390,L78,D287,L401,U638,R493,D493,R237,U29,R333,D466,L189,D634,R976,U934,R597,U62,L800,U229,R625,D927,L629,D334,L727,U4,R716,U20,L284,U655,R486,U883,R194,U49,L845,D960,R304,D811,L38,U114,R477,D318,L308,U445,L26,D44,R750,D12,R85,D146,R353,U715,R294,D595,L954,U267,L927,U383,L392,D866,L195,U487,L959,U630,R528,D482,R932,D541,L658,D171,L964,U687,R118,U53,L81,D381,R592,U238,L391,U399,R444,U921,R706,U925,R204,D220,L595,U489,R621,D15,R104,D567,L664,D54,R683,U654,R441,D748,L212,D147,L699,U296,L842,U230,L684,D470,R247,D421,R38,D757,L985,U791,R112,U494,R929,D726,L522,U381,R184,D492,L517,D819,R487,D620,R292,D206,R254,D175,L16,U924,R838,D423,R756,D720,L555,U449,L952,D610,L427,U391,R520,D957,R349,D670,L678,U467,R804,U757,L342,U808,R261,D597,L949,U162,R3,D712,L799,U531,R879,D355,R325,D173,L303,U679,L432,D421,R613,U431,L207,D669,R828,D685,R808,U494,R821,U195,L538,U16,L835,D442,L77,U309,L490,U918,L6,D200,L412,D272,L416,U774,L75,D23,L193,D574,R807,D382,L314,D885,R212,D806,L183,U345,L932,U58,L37,U471,R345,U678,R283,U644,L559,U892,L26,D358,L652,D606,L251,U791,L980,D718,L14,D367,R997,D812,R504,D474,L531,U708,R660,U253,L86,D491,R971,U608,L166,U659,R167,U92,L600,U20,R28,U852,R972,D409,L719,D634,R787,D796,L546,D857,L987,U111,L916,D108,R537,U931,R308,U385,L258,D171,R797,U641,R798,D723,R600,D710,R436,U587,R16,D564,L14,D651,R709,D587,R626,U270,R802,U937,R31,U518,L187,D738,R562,D238,R272,D403,R788,D859,L704,D621,L547,D737,L958,U311,L927"
cable_2 = "L1007,U199,L531,D379,L313,U768,L87,U879,R659,U307,L551,D964,L725,D393,R239,D454,R664,U402,R100,D62,R53,D503,R918,U998,L843,D142,R561,U461,R723,D915,L217,D126,L673,U171,R131,U518,R298,U99,L852,D799,L159,U161,R569,D802,L391,D553,L839,U954,R502,U351,R851,D11,L243,D774,L642,U613,R376,U867,L357,D887,L184,D298,R406,U912,R900,D348,L779,U494,R240,D712,R849,D684,R475,D449,L94,U630,L770,D426,L304,D869,R740,D377,R435,D719,L815,D211,R391,U484,R350,U599,R57,U210,R95,U500,L947,D649,R615,D404,L953,D468,R380,U215,R82,D872,R150,D347,L700,D450,L90,U803,L593,U296,R408,U667,R407,U725,R197,U269,R115,D758,R558,U62,L519,U308,R259,U869,L846,U3,R592,D357,R633,D909,L826,U165,L190,D821,L525,U404,R63,D214,R265,U74,L715,U461,L840,D647,R782,D655,R72,D601,L435,U549,L108,U687,R836,D975,L972,U813,R258,U703,R90,D438,R434,D818,R184,D886,R271,U31,L506,U395,L274,U455,R699,U889,L162,U771,R752,U317,R267,D959,R473,U58,R552,U51,R637,D726,R713,D530,L209,D824,R275,D207,R357,D373,L169,U880,L636,U409,R67,D740,R225,D272,R114,U970,R51,U230,R859,U820,L979,D153,R16,D883,L542,U806,L523,D752,L712,U55,L739,U746,R910,D873,R108,D558,L54,D619,L444,U941,R613,U478,L839,D910,R426,D614,L622,U928,L65,D644,L208,U971,L939,U874,R827,U672,L620,U140,L493,D637,L767,U831,R874,U468,R648,U714,R294,D606,L375,D962,L105,D919,L486,D771,L827,D196,L408,U217,L960,D633,L850,U805,L188,U566,L884,D212,L677,D204,R257,D409,R309,D783,L773,D588,L302,U129,L550,U919,L482,U563,R290,U690,R586,D460,L771,D63,R451,D591,L861,D598,L432,U818,L182"


def man_distance(cable_1, cable_2):
  directions_1 = cable_1.split(',')
  directions_2 = cable_2.split(',')
  
  path1 = {'x':[0],'y':[0], 'directions':directions_1}
  path2 = {'x':[0],'y':[0], 'directions':directions_2}
  
  paths = [path1, path2]

  for path in paths:
    i = 0
    for d in path['directions']:
      way = re.findall("^[A-Z]", d)[0]
      dist = int(re.findall("\d+", d)[0])
      for d in range(1,dist+1):
        if way == 'R':
          path["x"].append(path["x"][i] + 1)
          path["y"].append(path["y"][i])
        elif way == 'L':
          path["x"].append(path["x"][i] - 1)
          path["y"].append(path["y"][i])
        elif way == 'U':
          path["y"].append(path["y"][i] + 1)
          path["x"].append(path["x"][i])
        elif way == 'D':
          path["y"].append(path["y"][i] - 1)
          path["x"].append(path["x"][i])
        i = i + 1
    
  # print(path1)
  # print(path2)

  
  dif_len_y = len(path2['y']) - len(path1['y'])
  dif_len_x = len(path2['x']) - len(path1['x'])
  
  print(dif_len_y)
  print(path2['y'][-1])
  if dif_len_y > 0:
    path1['y'] = path1['y'] + [path1['y'][-1]] * dif_len_y
    path1['x'] = path1['x'] + [path1['x'][-1]] * dif_len_y
  if dif_len_y < 0:
    path2['y'] = path2['y'] + [path2['y'][-1]] * -dif_len_y
    path2['x'] = path2['x'] + [path2['x'][-1]] * -dif_len_x
  
  # print(path1['x'])
  # print(path2['x'])
  
  # x_index = [m - n for m,n in zip(path1['x'],path2['x'])]
  # y_index = [m - n for m,n in zip(path1['y'],path2['y'])]
  # 
  # print(x_index)
  # print(y_index)
  # 
  # print([i for i,val in enumerate(x_index) if val==0])
  # print([i for i,val in enumerate(y_index) if val==0])
  
  x_intersect = list(set(path1['x']) & set(path2['x']))
  y_intersect = list(set(path1['y']) & set(path2['y']))
  # print(x_intersect)
  # print(y_intersect)
  
  
  nodes = []
  
  for x in x_intersect:
    path1_x_indexes = [i for i,val in enumerate(path1['x']) if val==x]
    path2_x_indexes = [i for i,val in enumerate(path2['x']) if val==x]
    for i in path1_x_indexes:
      for j in path2_x_indexes:
        if path1['y'][i] == path2['y'][j]:
          nodes = nodes + [i]
        
  # print(nodes)
  
  # for l in loc:
  #   print(l)
  #   print(path1['x'][l])
  #   print(path2['x'][l])
  #   print(path1['y'][l])
  #   print(path2['y'][l])

  # 
  # 
  #   indexes = [i for i,val in enumerate(path1['x']) if val==x]
  #   # print(indexes)
  #   for i in indexes:
  #      & path1['x'][i] == path2['x'][i]:
  #       loc = loc + [i]
  # 
  # loc.remove(0)
  # print('potential closest nodes are:',loc)

  closest_node = None
  for node in nodes:
    maybe_closest = path1['x'][node] + path1['y'][node]
    if not closest_node or maybe_closest < closest_node:
      closest_node = maybe_closest
  print('closest node distance is:')
  print(closest_node)

  
  # [i for i,x in enumerate(set(path1['x']) if x == set(path2['x'])]
  # print(path1['x']))
  
  # for i in range(0,len(path1['x'])):
    # print(path2['x'][i])
  
  
man_distance(cable_1, cable_2)

```

Failed...

# --- Day 4: Secure Container ---

## Part 1

You arrive at the Venus fuel depot only to discover it's protected by a password. The Elves had written the password on a sticky note, but someone threw it out.

However, they do remember a few key facts about the password:

It is a six-digit number.
The value is within the range given in your puzzle input.
Two adjacent digits are the same (like 22 in 122345).
Going from left to right, the digits never decrease; they only ever increase or stay the same (like 111123 or 135679).
Other than the range rule, the following are true:

111111 meets these criteria (double 11, never decreases).
223450 does not meet these criteria (decreasing pair of digits 50).
123789 does not meet these criteria (no double).

**Problem: How many different passwords within the range given in your puzzle input meet these criteria?**



```{python}
def find_password(range_start, range_stop):
    final_potential_password = list()
    for maybe_password in range(range_start, range_stop+1):
        password_list = [int(i) for i in str(maybe_password)]
        code_ok = None
        if len(set(password_list)) < 6:
          code_ok = True
          previous_i = password_list[0]
          for i in password_list[1:]:
            if i < previous_i:
              code_ok = False
            previous_i = i
          if code_ok == True:
            #print('new potential password:' + str(maybe_password))
            final_potential_password.append(maybe_password)
    print('# of potential passwords: ' + str(len(final_potential_password)))
```

```{python}
test = [111111,223450,123789]
for i in test:
  find_password(i, i)
```

input range is : 125730-579381

```{python}
find_password(125730, 579381)
```

## Part 2

An Elf just remembered one more important detail: the two adjacent matching digits are not part of a larger group of matching digits.

Given this additional criterion, but still ignoring the range rule, the following are now true:

112233 meets these criteria because the digits never decrease and all repeated digits are exactly two digits long.
123444 no longer meets the criteria (the repeated 44 is part of a larger group of 444).
111122 meets the criteria (even though 1 is repeated more than twice, it still contains a double 22).

How many different passwords within the range given in your puzzle input meet all of the criteria?


```{python}
def find_password2(range_start, range_stop):
    final_potential_password = list()
    for maybe_password in range(range_start, range_stop+1):
        password_list = [int(j) for j in str(maybe_password)]
        pairs_ok = None
        previous_i = password_list[0]
        n = 1
        if len(set(password_list)) < 6:
          pairs_ok = True
          decrease_ok = True
          groups_ok = True
          for i in password_list[1:]:
              if i < previous_i :
                  decrease_ok = False
              if i == previous_i:
                n = n + 1
              elif i != previous_i and groups_ok == True:
                if n % 2 == 0:
                    groups_ok = True
                else:
                    groups_ok = False
                n = 1
              elif i != previous_i and groups_ok == False:
                n = 1
              previous_i = i
        if pairs_ok == True and decrease_ok == True and groups_ok == True:
            final_potential_password.append(maybe_password)
    print('# of potential passwords: ' + str(len(final_potential_password)))
```


```{python}
test = [112233,123444,111122]
for i in test:
  find_password2(i, i)
```
```{python}
find_password2(125730, 579381)
```


# --- Day 5: Sunny with a Chance of Asteroids ---

## Part 1

instruction can be found [here](https://adventofcode.com/2019/day/5)

**Problem: After providing 1 to the only input instruction and passing all the tests, what diagnostic code does the program produce?**

```{r}
intcode_computer <- function(gap) {
  op_index = 1
  gap_end = F
  op_index_list = c()
  
  while(gap_end == F){
      op_value <- gap[op_index]
      if (op_value == 1 ) {
        gap[gap[op_index + 3]+1] <- gap[gap[op_index + 1]+1] + gap[gap[op_index + 2]+1]
        op_index = op_index + 4
        # append(op_index_list,op_index)
      } else if (op_value == 2){
        gap[gap[op_index + 3]+1] <- gap[gap[op_index + 1]+1] * gap[gap[op_index + 2]+1]
        op_index = op_index + 4
        # append(op_index_list,op_index)
      } else if (op_value == 3) {
        op_index = op_index + 3
      } else if (op_value == 99) {
      gap_end = T
      return(gap)
      }
      print(op_value)
  }
}
```

```{r}
intcode_computer(c(1,0,0,0,99))
intcode_computer(c(2,3,0,3,99))
intcode_computer(c(1,1,1,4,99,5,6,0,99))
```

# 
